Quantifying social distancing behaviour across Australia in response to COVID-19

### 2020-04-19 manuscript/ms.Rmd
add notation to methods

 - General model description
   - latent factor model structure (weights)
   - construction of latent factors
   - observation model (weekends and holidays)
 - Data availability by source
 - Announcement timeline for Australia
 - Specific model fitting for Australia
    - interaction between distancing and weekends
    - greta HMC
    - residuals checking

>>>>

### General model description

#### Latent factor model

We construct a mechanistic latent-factor model that relates a series of underlying ('latent') population behaviours to an array of data streams, and then infer its parameters using statistical inference. This section, gives a general formulation of the model. Details of the specific implementation and data sources for Australia are given in a later section.

For each data stream $i$, we observe a time series of data values $\mathbf{y}_{i}$, comprising the values of the metric at each time $t$. We model the overall trend $\mathbf{m}_i$ in each of these data streams as a function of $K$ different temporal latent factors $\mathbf{z}_k$.

Each datastream trend is modelled as a linear function of the latent behavioural factors, with intercept term $\alpha_i$ and loading values $\omega_{i,k}$ against each latent factor:

$$
m_{i,t} = \alpha_i + \sum_{k=1}^K \omega_{i,k} z_{k,t}
$$
The latent factors $\mathbf{z}_k$ quantify general patterns of behaviour, rather than any one specific measurement of behaviour. As such, they are unitless, relative values which for statistical convenience we constrain to be scaled between 0 and 1. The sign and magnitude of the loadings $\omega_{i,k}$ therefore determine whether each behaviour corresponds to an increase or decrease in that metric, and by what magnitude.

#### Mechanistic latent factors

  We model each of these latent behavioural factors in a semi-mechanistic fashion. We compute the degree of behavioural change as a function of the timing, magnitude and early/late adoption of behavioural responses to events. For each latent factor $k$, we model the impact of each event $j$ as a function of: $\tau_{j,k}$ the time when the event took place; $\epsilon_{j,k}$ the relative impact of the event on behaviour; and $\kappa_{j,k}$, the variability among the population in how rapidly individuals responded to these events.

  For each latent behaviour $k$, these parameters describe the behavioural reponse to each event via a logistic distribution. Where the behaviour occurs at a single point in time, these parameters give rise to a logistic distribution over the times at which members of the population exhibit this behaviour, where the fraction of the population exhibiting behaviour $k$ in response to event $j$ at time $t$ is given by the probability density function of the logistic distribution:

$$
p_{j,k}(t) = \frac{e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}{\kappa_{j,k} (1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}) ^ 2}
$$

  Where the latent behavioural state is described by individuals switching from one behavioural state to another, the fraction of the population that have made the switch by time $t$ is given by the cumulative density function of the logistic distribution:

$$
p_{j,k}(t) = \frac{1}{1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}
$$

  Since the latent factors are scaled between 0 and 1, we scale these probabilities such that their maximum value is the relative event impact parameter $\epsilon_{j,k}$, and the sum over these parameeters is constrained to 1. Therefore, where the behaviour is a function of only a single event, $\epsilon_{j,k}$ is fixed at 1. The contribution of each event to the behaviour is therefore given by:

<!-- MAKE THIS HAVE DENSITY EPSILON AT THE MEAN -->
<!-- HAS DENSITY 1/(4 * kappa) AT MEAN -->
<!-- SO DIVIDE BY THAT AND MULTIPLY BY EPSILON -->
<!-- MUTIPLY BY epsilon * 4 * kappa -->

<!-- CHECK THIS! -->
$$
f_{j,k}(t) = 4 \kappa_{j,k} \epsilon_{j,k} \frac{e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}{\kappa_{j,k} (1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}) ^ 2}
$$
And the cumulative contribution by:

$$
f_{j,k}(t) = \frac{\epsilon_{j,k}}{1 + e^{-(t - \tau_{j,k}) / \kappa_{j,k}}}
$$

#### Data observatrion models

   - observation model (weekends, holidays, and interactions)

 - Data availability by source
 - Announcement timeline for Australia
 - Specific model fitting for Australia
    - interaction between distancing and weekends
    - greta HMC
    - residuals checking
